###Loading and preprocessing the data:
1 - Reading the data from the "activity.csv"
```{r, echo=TRUE}
setwd("/Users/Ali/Desktop/Pro")
activity <- NULL
activity <- read.csv("activity.csv", header = T, sep = ",")
```
2 - Summarizing:
```{r, echo=TRUE}
summary(activity)
```
```{r}
str(activity)
```
3 - Getting first few rows of the dataset:
```{r, echo=TRUE}
head(activity)
```
4 - Changing "date" variable to Date class;Changing "interval" variable to a factor:
```{r, echo=TRUE}
activity$date <- as.Date(activity$date, format = "%Y-%m-%d")
activity$interval <- factor(activity$interval)
```
###Calculating mean for total number of steps taken per day?
1 - Subsitting the dataset to ignore missing values
```{r, echo=TRUE}
NA_index <- is.na(as.character(activity$steps))
activity_no_NA <- activity[!NA_index,]
head(activity_no_NA)
```
2 - Aggregaton of the number of steps taken each day:
```{r, echo=TRUE}
#Creating a data frame with the steps taken for each day
steps_each_day <- aggregate(steps ~ date, data = activity_no_NA, sum)
#Adding column names to the created data frame
colnames(steps_each_day) <- c("date", "steps")
```
3 - Histogram of the total number of steps taken each day:
```{r, echo=TRUE}
hist(as.numeric(steps_each_day$steps), breaks = 20, col = "red", xlab = "Number of Steps", main= "Histogram of the total number of steps taken each day")
```
total number of steps taken per day:
```{r, echo=TRUE}
#Mean
mean(steps_each_day$steps)
```
```{r,echo=TRUE}
#Median
median(steps_each_day$steps)
```
###Analyzing the average daily activity pattern?
1 - Calculating the average number of steps taken, averaged across all days:
```{r, echo=TRUE}
#Calculating the average
steps_per_interval <- aggregate(activity_no_NA$steps, by=list(interval=activity_no_NA$interval), FUN=mean)

#Adding columns names
colnames(steps_per_interval) <- c("interval", "average_steps")

#ploting the average daily activity pattern 
plot(as.integer(levels(steps_per_interval$interval)), steps_per_interval$average_steps, type="l",
     xlab = "Interval", ylab = "Average Number of Steps", main = "Average Daily Activity Pattern",  col ="blue")
```
2 - The 5-minute interval that contains the maximum number of steps:
```{r, echo=TRUE}
#The maximum number of average steps
max_steps <- max(steps_per_interval$average_steps)
max_steps
```
```{r, echo=TRUE}
#The 5-minute interval that contains the maximum number of steps
intervale_max_steps<-steps_per_interval[which.max(steps_per_interval$average_steps),]$interval
intervale_max_steps
```
So, the 5-minute interval that contains the maximum number of steps ( 206.2 steps ) is the interval 835.

Imputing missing values
1 - The total number of missing values in the dataset (for each variable):

For the "steps" variable:
```{r, echo=TRUE}
sum(is.na(as.character(activity$steps)))
```
For the "interval" variable:
```{r, echo=TRUE}
sum(is.na(as.character(activity$interval)))
```
So, the total number of missing values in the dataset is 2304.





